Assimilating GOES-16

### 2022-05-03 analysis/chapters/DA_goes16.Rmd
more organization

library(mesoda)
library(metR)
library(tidyverse)
library(lubridate)
library(data.table)
library(here)
library(patchwork)
library(unglue)
library(knitr)
library(kableExtra)
source(here::here("temp_R", "goes_functions.R"))

@@

```{r}
nc <- ncdf4::nc_open('~/datosmunin3/DATA/GOES16/ABI-L2-ACMF/OR_ABI-L2-ACMF-M3_G16_s20183250000341_e20183250011108_c20183250011312.nc')

cld_msk <- ReadNetCDF('~/datosmunin3/DATA/GOES16/ABI-L2-ACMF/OR_ABI-L2-ACMF-M3_G16_s20183250000341_e20183250011108_c20183250011312.nc', 
                      vars = 'BCM', 
                      subset = list(x = xlim,
                                    y = ylim)) %>% 
  .[, c("lon", "lat") := goes_projection(x, y, nc)]

cld_msk %>% 
  ggplot(aes(lon, lat)) +
  geom_point(aes(color = BCM == 0)) +
  scale_color_manual(values = c("darkgrey", "white"), guide = NULL) +
  geom_mapa() +
  labs(x = NULL, y = NULL,
       title = "Cloud Mask", subtitle = "Time: 2018-11-21 00 UTC")

@@

```{r}
files <- Sys.glob("~/datosmunin3/EXP/prueba_abi/ANA/*00/diagfiles/asim_abi_g16_*.ensmean")

@@

  # meta <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/prueba_abi/ANA/20181121000000/diagfiles/asim_abi_g16_20181121000000.ensmean")

  read_diag_rad(f, "EG1") %>% 
    # .[, run := meta[[1]][["exp"]]] %>% 
    .[, channel := channel + 6]

@@

@@

diag %>% 
  .[channel %in% c(8:10)] %>% 
  .[, date := format(date, "%m/%d %H UTC")] %>%
  ggplot(aes(factor(channel))) +
  geom_bar(aes(fill = factor(qc))) +
  scale_fill_manual(values = c("darkorange","grey40", "grey60", "grey80"),
                    labels = c("0" = "pass qc", 
                               "3" = "gross check",
                               "5" = "surface check",
                               "50" = "std dev")) +
  facet_wrap(~date, ncol = 7) +
  labs(fill = NULL, x = "Channel") +
  theme_minimal() +
  theme(legend.position = "bottom")

@@

@@

```{r}
diag %>% 
  .[channel %in% c(10)] %>% 
  .[, date := format(date, "%m/%d %H UTC")] %>%
  ggplot(aes(ConvertLongitude(lon), lat)) +
  geom_point(aes(color = factor(qc)), size = 0.9) +
  geom_mapa() +
  scale_color_manual(values = c("grey50", "grey70", "grey80", "cyan4"),
                     labels = c("0" = "pass qc", 
                                "3" = "gross check",
                                "5" = "surface check",
                                "50" = "std dev")) +
  guides(color = guide_legend(override.aes = list(size = 2))) +
  facet_wrap(~date, ncol = 4) +
  labs(subtitle = "Channel 10", x =  NULL, y = NULL, color = "QC") +
  theme_minimal() +
  theme(legend.position = "bottom")
```
Only a few observations are rejected at each cycle. These observations are located over The Andes (surface type = snow)

@@

```{r}
files_ana <- Sys.glob("~/datosmunin3/EXP/prueba_abi/ANA/*00/analysis.mem001")

@@

  meta <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/prueba_abi/ANA/{date}/analysis.mem001")

@@

files_gue <- Sys.glob("~/datosmunin3/EXP/prueba_abi/GUESS/*00/wrfarw.mem001")

@@

  meta <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/prueba_abi/GUESS/{date}/wrfarw.mem001")

@@

@@

  ggsave(paste0("t_ab_", format(dates[i], "%Y-%m-%d_%H"), ".png"), height = 3)

@@

  ggsave(paste0("q_ab_", format(dates[i], "%Y-%m-%d_%H"), ".png"), height = 3)

@@

  ggsave(paste0("t_ab_", levs[i], ".png"), height = 4)

@@

  ggsave(paste0("q_ab_", levs[i], ".png"), height = 4)

@@

```
```{r}
files <- Sys.glob("~/datosmunin3/EXP/prueba_abi/ANA/*00/diagfiles/asim_conv*.ensmean")

>>>>

library(mesoda)
library(metR)
library(tidyverse)
library(lubridate)
library(data.table)
library(here)
library(patchwork)
library(unglue)
library(knitr)
library(kableExtra)
# source(here::here("temp_R", "goes_functions.R"))

@@

```{r}
  nc <- ncdf4::nc_open('~/datosmunin3/DATA/GOES16/ABI-L2-ACMF/OR_ABI-L2-ACMF-M3_G16_s20183240000342_e20183240011109_c20183240011270.nc')

cld_msk <- ReadNetCDF('~/datosmunin3/DATA/GOES16/ABI-L2-ACMF/OR_ABI-L2-ACMF-M3_G16_s20183240000342_e20183240011109_c20183240011270.nc', 
                      vars = 'BCM', 
                      subset = list(x = xlim,
                                    y = ylim)) %>% 
  .[, c("lon", "lat") := goes_projection(x, y, nc)]

cld_msk %>% 
  ggplot(aes(lon, lat)) +
  geom_point(aes(color = BCM == 0)) +
  scale_color_manual(values = c("darkgrey", "white"), guide = NULL) +
  geom_mapa() +
  labs(x = NULL, y = NULL,
       title = "Cloud Mask", subtitle = "Time: 2018-11-20 00 UTC")

@@

```{r}
files <- Sys.glob("~/datosmunin3/EXP/E10/ANA/*00/diagfiles/asim_abi_g16_*.ensmean")

@@

  read_diag_rad(f, "E10") %>% 
    .[, channel := channel + 6]

files <- Sys.glob("~/datosmunin3/EXP/EG2/ANA/*00/diagfiles/asim_abi_g16_*.ensmean")

diag_eg2 <- map(files, function(f) {

  read_diag_rad(f, "EG2") %>% 
    .[, channel := channel + 6]

@@

```{r}
rbind(diag, diag_eg2) %>% 
  .[channel %in% c(8:10), .N, by = .(qc, date, channel, exp)] %>% 
  ggplot(aes(date, N)) +
  geom_point(aes(color = factor(qc))) +
  geom_line(aes(color = factor(qc))) +
  # scale_color_manual(values = c("darkorange","grey40", "grey60", "grey80"),
  #                    labels = c("0" = "pass qc", 
  #                               "3" = "gross check",
  #                               "5" = "surface check",
  #                               "50" = "std dev")) +
  labs(x = NULL, y = NULL, color = NULL) +
  facet_grid(exp~channel,  scales = "free_y",
             labeller = labeller(channel = channels)) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

@@

diag %>% 
  .[channel %in% c(8:10)] %>% 
  .[, date := format(date, "%m/%d %H UTC")] %>%
  ggplot(aes(factor(channel))) +
  geom_bar(aes(fill = factor(qc))) +
  scale_fill_manual(values = c("darkorange","grey40", "grey60", "grey80"),
                    labels = c("0" = "pass qc", 
                               "3" = "gross check",
                               "5" = "surface check",
                               "50" = "std dev")) +
  facet_wrap(~date, ncol = 7) +
  labs(fill = NULL, x = "Channel", subtitle = "EG1") +
  theme_minimal() +
  theme(legend.position = "bottom")

diag_eg2 %>% 
  .[channel %in% c(8:10)] %>% 
  .[, date := format(date, "%m/%d %H UTC")] %>%
  ggplot(aes(factor(channel))) +
  geom_bar(aes(fill = factor(qc))) +
  scale_fill_manual(values = c("darkorange","grey40", "grey60", "grey80"),
                    labels = c("0" = "pass qc", 
                               "3" = "gross check",
                               "5" = "surface check",
                               "50" = "std dev")) +
  facet_wrap(~date, ncol = 7) +
  labs(fill = NULL, x = "Channel", subtitle = "EG2") +
  theme_minimal() +
  theme(legend.position = "bottom")

@@

```{r}
diag_eg2 %>% 
  .[channel %in% c(8:10)] %>% 
  .[, date := format(date, "%m/%d %H UTC")] %>%
  ggplot(aes(ConvertLongitude(lon), lat)) +
  geom_point(aes(color = factor(qc)), size = 0.9) +
  geom_mapa() +
  scale_color_manual(values = c("darkorange","grey40", "grey60", "grey80"),
                     labels = c("0" = "pass qc", 
                                "3" = "gross check",
                                "5" = "surface check",
                                "50" = "std dev")) +
  guides(color = guide_legend(override.aes = list(size = 2))) +
  facet_grid(channel~date) +
  labs(subtitle = "Assimilated obs (EG2)", x =  NULL, y = NULL, color = "QC") +
  theme_minimal() +
  theme(legend.position = "bottom")

@@

```{r}
diag %>% 
  .[channel %in% c(10)] %>% 
  .[, date := format(date, "%m/%d %H UTC")] %>%
  ggplot(aes(ConvertLongitude(lon), lat)) +
  geom_point(aes(color = factor(qc)), size = 0.9) +
  geom_mapa() +
  scale_color_manual(values = c("grey50", "grey70", "grey80", "cyan4"),
                     labels = c("0" = "pass qc", 
                                "3" = "gross check",
                                "5" = "surface check",
                                "50" = "std dev")) +
  guides(color = guide_legend(override.aes = list(size = 2))) +
  facet_wrap(~date, ncol = 4) +
  labs(subtitle = "Channel 10", x =  NULL, y = NULL, color = "QC") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
diag_eg2 %>% 
  .[channel %in% c(10)] %>% 
  .[, date := format(date, "%m/%d %H UTC")] %>%
  ggplot(aes(ConvertLongitude(lon), lat)) +
  geom_point(aes(color = factor(qc)), size = 0.9) +
  geom_mapa() +
  scale_color_manual(values = c("grey50", "grey70", "grey80", "cyan4"),
                     labels = c("0" = "pass qc", 
                                "3" = "gross check",
                                "5" = "surface check",
                                "50" = "std dev")) +
  guides(color = guide_legend(override.aes = list(size = 2))) +
  facet_wrap(~date, ncol = 4) +
  labs(subtitle = "Channel 10 (EG2)", x =  NULL, y = NULL, color = "QC") +
  theme_minimal() +
  theme(legend.position = "bottom")
```
Only a few observations are rejected at each cycle. These observations are located over The Andes (surface type = snow)

@@

```{r}
files_ana <- Sys.glob("~/datosmunin3/EXP/E9/ANA/*00/analysis.mem001")[c(1:7)]

@@

  meta <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/E6/ANA/{date}/analysis.mem001")

@@

files_gue <- Sys.glob("~/datosmunin3/EXP/EG2/GUESS/*00/wrfarw.mem001")

@@

  meta <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/EG2/GUESS/{date}/wrfarw.mem001")

@@

files_conv <- Sys.glob("~/datosmunin3/EXP/E2/ANA/*00/analysis.mem001")[1:7]

ana_conv <- map(files_conv, function(f) {

  meta <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/E2/ANA/{date}/analysis.mem001")

  ReadNetCDF(f, vars = c(t_conv = "T", p = "P", "PB", q_conv = "QVAPOR",
                         lon = "XLONG", lat = "XLAT")) %>% 
    .[, ":="(t_conv = tk(t_conv, p + PB, T_BASE = 290) - 273.15,
             p = p + PB,
             PB = NULL)] %>% 
    .[, date := ymd_hms(meta[[1]][["date"]])] %>% 
    .[, .(lev = levs,
          t_conv = approx(p, t_conv, xout = levs*100)$y,
          q_conv = approx(p, q_conv, xout = levs*100)$y),
      by = .(date, south_north, west_east)] %>%
    coord[., on = .NATURAL] %>% 
    .[, c("x", "y") := wrf_project(lon, lat)] %>% 
    .[]

files_conv <- Sys.glob("~/datosmunin3/EXP/*/ANA/20181121000000/analysis.ensmean")[c(1:5,7:8)]

ana_conv <- map(files_conv, function(f) {

  meta <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/{run}/ANA/20181121000000/analysis.ensmean")

  ReadNetCDF(f, vars = c(t_conv = "T", p = "P", "PB", q_conv = "QVAPOR",
                         lon = "XLONG", lat = "XLAT")) %>% 
    .[, ":="(t_conv = tk(t_conv, p + PB, T_BASE = 290) - 273.15,
             p = p + PB,
             PB = NULL)] %>% 
    .[, date := meta[[1]][["run"]]] %>% 
    .[, .(lev = levs,
          t_conv = approx(p, t_conv, xout = levs*100)$y,
          q_conv = approx(p, q_conv, xout = levs*100)$y),
      by = .(date, south_north, west_east)] %>%
    coord[., on = .NATURAL] %>% 
    .[, c("x", "y") := wrf_project(lon, lat)] %>% 
    .[]

}) %>% 
  rbindlist()
```

```{r}
temp <- ana_conv %>% 
  dcast(x + y + lev ~ date, value.var = "q_conv") %>% 
  .[, ":="(aut = E5 - E2,
           satwnd = E6 - E5,
           rad_bad = E7 - E6,
           rad = E8 - E6,
           # abi_old = prueba_abi - E2,
           abi = EG2 - E6,
           abi_sat = EG2 - E5,
           abi_sat_aut = EG2 - E2)] %>% 
  melt(measure.vars = c("aut", "satwnd", "rad_bad", "rad", "abi", "abi_sat", "abi_sat_aut"))

temp2 <- ana_conv %>% 
  dcast(x + y + lev ~ date, value.var = "t_conv") %>% 
  .[, ":="(aut = E5 - E2,
           satwnd = E6 - E5,
           rad_bad = E7 - E6,
           rad = E8 - E6,
           # abi_old = prueba_abi - E2,
           abi = EG2 - E6,
           abi_sat = EG2 - E5,
           abi_sat_aut = EG2 - E2)] %>% 
  melt(measure.vars = c("aut", "satwnd", "rad_bad", "rad", "abi", "abi_sat", "abi_sat_aut"))

for (i in 1:length(levs)) {
  temp %>% 
    .[lev == levs[i]] %>% 
    ggplot(aes(x, y)) +
    geom_contour_fill(aes(z = value*1000, fill = ..level..), proj = norargentina_lambert) +
    scale_fill_divergent_discretised(trans = mesoda::symlog_trans(),
                                     guide = guide_colorsteps(barhight = 10)) +
    geom_mapa() +
    labs(x = NULL, y = NULL, fill = NULL,
         title = "Humidity",
         subtitle = as.character(levs[i])) +
    facet_wrap(~variable, ncol = 4) +
    theme_minimal()

   ggsave(paste0("impact_2018112100_q_", levs[i], ".png"), height = 4)

  temp2 %>% 
    .[lev == levs[i]] %>% 
    ggplot(aes(x, y)) +
    geom_contour_fill(aes(z = value, fill = ..level..), proj = norargentina_lambert) +
    scale_fill_divergent_discretised(trans = mesoda::symlog_trans(),
                                     guide = guide_colorsteps(barhight = 10)) +
    geom_mapa() +
    labs(x = NULL, y = NULL, fill = NULL,
         title = "Temperature",
         subtitle = as.character(levs[i])) +
    facet_wrap(~variable, ncol = 4) +
    theme_minimal()

   ggsave(paste0("impact_2018112100t_", levs[i], ".png"), height = 4)

@@

  ggsave(paste0("EG2_t_ab_", format(dates[i], "%Y-%m-%d_%H"), ".png"), height = 3)

@@

  ggsave(paste0("EG2_q_ab_", format(dates[i], "%Y-%m-%d_%H"), ".png"), height = 3)

@@

  ggsave(paste0("EG2_t_ab_", levs[i], ".png"), height = 4)

@@

  ggsave(paste0("EG2_q_ab_", levs[i], ".png"), height = 4)

@@

```{r}
for (i in 1:length(levs)) {

  ana[ana_conv, on = .NATURAL] %>% 
    .[lev == levs[i]] %>% 
    .[, date := format(date, "%m/%d %H UTC")] %>%
    .[, a_g := t_ana - t_conv] %>%
    ggplot(aes(x, y)) +
    geom_contour_fill(aes(z = a_g, fill = ..level..), proj = norargentina_lambert) +
    scale_fill_divergent_discretised() +
    geom_mapa() +
    labs(x = NULL, y = NULL, fill = "SATWND-CONV",
         title = "Temperature",
         subtitle = paste(as.character(levs[i]), "hPa")) +
    facet_wrap(~date, ncol = 4) +
    theme_minimal()

  ggsave(paste0("E62_t_ac_", levs[i], ".png"), height = 4)

  ana[ana_conv, on = .NATURAL] %>% 
    .[lev == levs[i]] %>% 
    .[, date := format(date, "%m/%d %H UTC")] %>%
    .[, a_g := q_ana - q_conv] %>%
    ggplot(aes(x, y)) +
    geom_contour_fill(aes(z = a_g*1000, fill = ..level..), proj = norargentina_lambert) +
    scale_fill_divergent_discretised() +
    geom_mapa() +
    labs(x = NULL, y = NULL, fill = "SATWND-CONV",
         title = "Humidity",
         subtitle = paste(as.character(levs[i]), "hPa")) +
    facet_wrap(~date, ncol = 4) +
    theme_minimal()

  ggsave(paste0("E62_q_ac_", levs[i], ".png"), height = 4)

}
```
```{r}
files_ana <- Sys.glob("~/datosmunin3/EXP/prueba_abi/ANA/*_*/analysis.mem001")[c(1,3:5)]

  meta <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/prueba_abi/ANA/20181121000000_{run}/analysis.mem001")

  ReadNetCDF(f, vars = c(t_ana = "T", p = "P", "PB", q_ana = "QVAPOR",
                         lon = "XLONG", lat = "XLAT")) %>% 
    .[, ":="(t_ana = tk(t_ana, p + PB, T_BASE = 290) - 273.15,
             p = p + PB,
             PB = NULL)] %>% 
    .[, run := meta[[1]][["run"]]] %>% 
    .[, .(lev = levs,
          t_ana = approx(p, t_ana, xout = levs*100)$y,
          q_ana = approx(p, q_ana, xout = levs*100)$y),
      by = .(run, south_north, west_east)] %>%
    coord[., on = .NATURAL] %>% 
    .[, c("x", "y") := wrf_project(lon, lat)] %>% 
    .[]

ana %>% 
  .[lev == levs[i]] %>% 
  # .[, date := format(date, "%m/%d %H UTC")] %>%
  # .[, a_g := q_ana - q_conv] %>%
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = q_ana*1000, fill = ..level..), proj = norargentina_lambert) +
  # scale_fill_divergent_discretised() +
  geom_mapa() +
  labs(x = NULL, y = NULL, fill = NULL,
       title = "Humidity",
       subtitle = paste(as.character(levs[i]), "hPa")) +
  facet_wrap(~run, ncol = 4) +
  theme_minimal()

ana[run != "all"] %>% 
  dcast(x + y + lev ~ run, value.var = "q_ana") %>% 
  .[ana[run == "all"], on = .NATURAL] %>% 
  .[, ":="(all_8 = t_ana - ch08,
           all_9 = t_ana - ch09,
           all_0 = t_ana - ch10)] %>% 
  melt(measure.vars = c("all_8", "all_9", "all_0")) %>% 
  .[lev == levs[4]] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = value, fill = ..level..), proj = norargentina_lambert) +
  scale_fill_divergent_discretised() +
  geom_mapa() +
  labs(x = NULL, y = NULL, fill = NULL,
       title = "Humidity",
       subtitle = paste(as.character(levs[i]), "hPa")) +
  facet_wrap(~variable, ncol = 4) +
  theme_minimal()

```{r}
files <- Sys.glob("~/datosmunin3/EXP/prueba_abi/ANA/*00/diagfiles/asim_conv*.ensmean")
